---
title: 'RQ1 - Trends in mental health care and telehealth use across area deprivation: An analysis of electronic health records from 2016 to 2024'
author: Grace Ringlein
date: "01-31-2025"
output:
  html_document: 
    df_print: paged
  html_notebook:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(lubridate)
library(marginaleffects)
library(here)
library(gtsummary)
library(readxl)
```


```{r helpers, include=FALSE}
'%notin%' <- Negate('%in%')

get_num_pats <- function(data) {
  data %>% pull(cohort_id) %>% unique() %>% length()
}

get_mode <- function(x) {
  uniqx <- unique(x)
  uniqx[which.max(tabulate(match(x, uniqx)))]
}


relabel_program <- function(data) {
  data %>% mutate(program= factor(ifelse(program=="PSYCH","Mental health care in psychiatry","Mental health care in primary care"),levels=c("Mental health care in primary care","Mental health care in psychiatry")))
}

relabel_program_v2 <- function(data) {
  data %>% mutate(program= factor(ifelse(program=="PSYCH","Mental health care\n in psychiatry","Mental health care\n in primary care"),levels=c("Mental health care\n in primary care","Mental health care\n in psychiatry")))
}

```

```{r load_data}
data <- readr::read_rds(here("telehealth","data","tele_samples11062024_from2016.rds")) 
```

```{r}
top_five_county <- rbind(data$tele_psych_F,data$tele_primary_F) %>% select(contains("county"),cohort_id) %>% unique() %>% count(county) %>% arrange(desc(n)) %>% head(5) %>% pull(county)


reg_enc_psych <-data$tele_psych_F %>% 
  arrange(cohort_id) %>%
  group_by(cohort_id) %>% 
  arrange(contact_date,.by_group=TRUE) %>%
  ungroup()  %>%  
  mutate(mode_payor_cat = ifelse(is.na(mode_payor_cat), "Other",mode_payor_cat),
         county_top = 
           case_when(
             county %in% top_five_county ~ county,
             TRUE ~ "Other"))  %>%
  mutate(ADI_group =
           factor(case_when(
             adi_nat <= 25 ~ "Low deprivation",
             adi_nat <= 75 ~ "Medium deprivation",
             adi_nat <= 100 ~ "High deprivation",
             ),levels=c("High deprivation","Medium deprivation","Low deprivation")),
         time_point = factor(case_when(
           contact_date < as_date("2019-07-01") ~ "Pre-pandemic",
           contact_date < as_date("2020-07-01") ~ "Emerging-pandemic",
           contact_date < as_date("2021-07-01") ~ "Early-pandemic",
           contact_date < as_date("2024-07-01") ~ "Post-pandemic"),
           levels = c("Pre-pandemic","Emerging-pandemic", "Early-pandemic","Post-pandemic"))) %>% 
  group_by(cohort_id) %>% 
  mutate(
      in_pre = sum(time_point == "Pre-pandemic") > 0,
      in_emerging = sum(time_point == "Emerging-pandemic") > 0,
      in_early = sum(time_point == "Early-pandemic") > 0,
      in_post = sum(time_point == "Post-pandemic") > 0,
      in_pre_post = in_pre & in_post) %>% 
  ungroup()
         
  
reg_enc_primary <- data$tele_primary_F %>% 
  arrange(cohort_id) %>%
  group_by(cohort_id) %>% 
  arrange(contact_date,.by_group=TRUE) %>%
  ungroup()  %>%  
  mutate(mode_payor_cat = ifelse(is.na(mode_payor_cat), "Other",mode_payor_cat),
         county_top = 
           case_when(
             county %in% top_five_county ~ county,
             TRUE ~ "Other"))  %>%
  mutate(ADI_group =
           factor(case_when(
             adi_nat <= 25 ~ "Low deprivation",
             adi_nat <= 75 ~ "Medium deprivation",
             adi_nat <= 100 ~ "High deprivation",
             ),levels=c("High deprivation","Medium deprivation","Low deprivation")),
         time_point = factor(case_when(
           contact_date < as_date("2019-07-01") ~ "Pre-pandemic",
           contact_date < as_date("2020-07-01") ~ "Emerging-pandemic",
           contact_date < as_date("2021-07-01") ~ "Early-pandemic",
           contact_date < as_date("2024-07-01") ~ "Post-pandemic"),
           levels = c("Pre-pandemic","Emerging-pandemic", "Early-pandemic","Post-pandemic"))) %>% 
  group_by(cohort_id) %>% 
  mutate(
      in_pre = sum(time_point == "Pre-pandemic") > 0,
      in_emerging = sum(time_point == "Emerging-pandemic") > 0,
      in_early = sum(time_point == "Early-pandemic") > 0,
      in_post = sum(time_point == "Post-pandemic") > 0,
      in_pre_post = in_pre & in_post) %>% 
  ungroup()
         
reg_enc_combo <- rbind(reg_enc_psych,reg_enc_primary)

reg_enc_psych <- reg_enc_psych %>%
  mutate(in_both_cohorts = cohort_id %in% reg_enc_primary$cohort_id) %>% 
  group_by(cohort_id) %>%
  mutate(n_visits=n()) %>% 
  ungroup()

reg_enc_primary <- reg_enc_primary %>%
  mutate(in_both_cohorts = cohort_id %in% reg_enc_psych$cohort_id)%>%
  group_by(cohort_id) %>%
  mutate(n_visits=n()) %>%
  ungroup()
```

## How many census block groups? 

### Number of census block groups in the combined primary care + psych 2016-2024 samples:

Overall: 
```{r}
reg_enc_combo %>% count(Geoid_2020) %>% nrow()
```

By ADI group: 
```{r}
reg_enc_combo %>% group_by(ADI_group) %>% summarize(n_block_groups=length(unique(Geoid_2020))) 
```

### Number of census block groups in the primary care 2016-2024 sample:

Overall: 
```{r}
reg_enc_primary %>% count(Geoid_2020) %>% nrow() %>% print()
```

By ADI group: 
```{r}
reg_enc_primary %>% group_by(ADI_group) %>% summarize(n_block_groups=length(unique(Geoid_2020))) 
```

### Number of census block groups in the psych 2016-2024 sample:

Overall:
```{r}
reg_enc_psych %>% count(Geoid_2020) %>% nrow()
```

By ADI group: 
```{r}
reg_enc_psych %>% group_by(ADI_group) %>% summarize(n_block_groups=length(unique(Geoid_2020))) 

```

## Distribution of ADI among patients
```{r adi_dist}
reg_enc_combo %>% select(cohort_id,adi_nat,ADI_group,program) %>% unique() %>% relabel_program() %>% ggplot() + geom_histogram(aes(x=adi_nat,fill=ADI_group),bins=100) +
  xlab("ADI (National Percentile) 2020") + theme_bw() + facet_wrap(.~program) + 
   scale_fill_manual(name="ADI Group",breaks=c("Low deprivation","Medium deprivation","High deprivation"),
                    values=c("#FABF7B","#E05C5C","#6E005F"))+  theme(legend.position=c(.85,.82),
        legend.background = element_rect(colour = "black")) + 
  ylab("N")

ggsave(here("figures","fig_S3_2016-2024cohort.png"))
```

```{r}
# get average number of appointments per group per day 
tmp <- rbind(reg_enc_primary,reg_enc_psych) %>% 
  group_by(contact_date,program,month,year,ADI_group,time_point, t_months) %>% filter(!weekend) %>%
  summarize(n=n()) %>%
  group_by(program,month,year,ADI_group,time_point, t_months) %>%
  summarize(first_date = min(contact_date),
           n = mean(n)) %>% ungroup()


# average numbers per day in july 2018- july 2019 for comparison 
tmp_base <- rbind(reg_enc_primary,reg_enc_psych) %>% 
  filter(contact_date >= as_date("2018-07-01") & contact_date < as_date("2019-07-01")) %>% 
  group_by(contact_date,program,month,year,ADI_group, t_months) %>% 
  summarize(n=n()) %>% 
  group_by(ADI_group,program) %>%
  summarize(
           n_base = mean(n)) %>% ungroup()

# join 
tmp <- tmp %>% left_join(tmp_base,by=c("ADI_group","program")) %>% group_by(ADI_group,program) %>%
  mutate(n_perc = (n/n_base - 1)* 100) %>% ungroup()
```


```{r}
# plot number of visits per day for each group (extra figure, not in paper)
tmp %>% relabel_program() %>% ggplot(aes(x=first_date,y=n,colour=ADI_group)) + geom_point() + geom_line() +  facet_grid(rows=vars(program)) + 
   geom_vline(xintercept=as_date("2020-03-01"),linetype="dashed") + 
   scale_x_continuous(breaks=c(as_date("2016-07-01"),as_date("2017-07-01"),as_date("2018-07-01"),as_date("2019-07-01"),as_date("2020-07-01"),as_date("2021-07-01"),as_date("2022-07-01"),as_date("2023-07-01"),as_date("2024-07-01")),
                     labels = paste("July",2016:2024)) + 
   scale_colour_manual(name="ADI Group",breaks=c("Low deprivation","Medium deprivation","High deprivation"),
                    values=c("#FABF7B","#E05C5C","#6E005F"))+
  theme_bw() +   theme(legend.position=c(.15,.9),axis.text.x = element_text(angle = 60, hjust=1),
        legend.background = element_rect(colour = "black"),
        legend.text = element_text(size=8),
        legend.title = element_text(size=8),
        legend.key.size = unit(.1, "cm")) + 
  ylim(c(0,150)) + 
  ylab("Visits per day") + xlab("")
ggsave(here("figures","extra_fig_volume_of_care.png"))
```


```{r}
# plot relative number of visits per day for each group 
tmp %>% relabel_program_v2() %>% ggplot() +
  geom_hline(yintercept=0,alpha=.3) + 
  geom_rect(xmin=as_date("2019-07-01"),xmax=as_date("2021-07-01"),ymin=-Inf,ymax=Inf,fill="#EEEEEE",alpha=.3) + 
   geom_vline(xintercept=as_date("2019-07-01"),linetype="dotted") + 
  geom_vline(xintercept=as_date("2020-03-01"),linetype="dashed") + 
  geom_vline(xintercept=as_date("2021-07-01"),linetype="dotted") + 
  scale_x_continuous(breaks=c(as_date("2016-07-01"),as_date("2017-07-01"),as_date("2018-07-01"),as_date("2019-07-01"),as_date("2020-07-01"),as_date("2021-07-01"),as_date("2022-07-01"),as_date("2023-07-01"),as_date("2024-07-01")),
                     labels = paste("July",2016:2024)) + 
  geom_point(aes(x=first_date,y=n_perc,colour=ADI_group)) + geom_line(aes(x=first_date,y=n_perc,colour=ADI_group)) + facet_grid(rows=vars(program)) + 
   scale_colour_manual(name="ADI Group",breaks=c("Low deprivation","Medium deprivation","High deprivation"),
                    values=c("#FABF7B","#E05C5C","#6E005F"))+
  theme_bw() +   theme(legend.position=c(.15,.9), axis.text.x = element_text(angle = 60, hjust=1),
       legend.background = element_rect(colour = "black"),
        legend.text = element_text(size=8),
        legend.title = element_text(size=8),
        legend.key.size = unit(.1, "cm")) + 
  
  #ylim(c(0,)) + 
  ylab("% change daily visits\n relative to July 2018-June 2019 average") + xlab("") 

ggsave(here("figures","fig_1.png"),height=6,width=7)

```


## Figure S6 Trends in mental health utilization 

```{r}
# data organized for plots 
tmp_analysis_nogroup <- rbind(reg_enc_primary,reg_enc_psych) %>% 
  group_by(contact_date,program,month,year,time_point, t_months,day) %>% 
  summarize(n=n(),
            tele = mean(telehealth),
            hosp=mean(hosp)) %>%
  group_by(program,t_months) %>% 
  summarize(first_date = min(contact_date),
           n = mean(n),
           hosp=mean(hosp),
           tele=mean(tele) * 100) %>% ungroup()
```


```{r}
# panel 1 with legend
tmp_analysis_nogroup %>% 
  relabel_program() %>% ggplot() +
  scale_x_continuous(breaks=c(as_date("2016-07-01"),as_date("2017-07-01"),as_date("2018-07-01"),as_date("2019-07-01"),as_date("2020-07-01"),as_date("2021-07-01"),as_date("2022-07-01"),as_date("2023-07-01"),as_date("2024-07-01")),
                     labels = paste("July",2016:2024)) + 
  geom_vline(xintercept=as_date("2020-03-01"),linetype="solid") + 
  geom_vline(xintercept=as_date("2021-07-01"),linetype="dashed") + 
  geom_vline(xintercept=as_date("2021-12-01"),linetype="dotted") + 
  geom_vline(xintercept=as_date("2023-05-01"),linetype="dotdash") + 
  geom_point(aes(x=first_date,y=n,colour=program)) +
  geom_line(aes(x=first_date,y=n,colour=program)) + 
  theme_bw() +   
  scale_colour_manual(name="Department",values=c("#FFA47F","#1E99E5")) + 
  ylab("Daily mental health care visits") + xlab("")+   theme(legend.position=c(.2,.75),
        legend.background = element_rect(colour = "black"),
        axis.text.x = element_text(angle = 60, hjust=1),
        legend.text = element_text(size=8),
        legend.title = element_text(size=8),
        axis.title.y = element_text(size=8),
        legend.key.size = unit(.1, "cm"))

ggsave(here("figures","figure_S6_p1_wlegend.png"),width=7,height=2.5)
```


```{r}
# panel 1 with/o legend
tmp_analysis_nogroup %>% 
  relabel_program() %>% ggplot() +
  geom_vline(xintercept=as_date("2020-03-01"),linetype="solid") + 
  geom_vline(xintercept=as_date("2021-07-01"),linetype="dashed") + 
  geom_vline(xintercept=as_date("2021-12-01"),linetype="dotted") + 
  geom_vline(xintercept=as_date("2023-05-01"),linetype="dotdash") + 
  scale_x_continuous(breaks=c(as_date("2016-07-01"),as_date("2017-07-01"),as_date("2018-07-01"),as_date("2019-07-01"),as_date("2020-07-01"),as_date("2021-07-01"),as_date("2022-07-01"),as_date("2023-07-01"),as_date("2024-07-01")),
                     labels = paste("July",2016:2024)) + 

  geom_point(aes(x=first_date,y=n,colour=program)) +
  geom_line(aes(x=first_date,y=n,colour=program)) + 
  theme_bw() +   
  scale_colour_manual(name="Department",values=c("#FFA47F","#1E99E5")) + 
  ylab("Daily mental health care visits") + xlab("")+   
  theme(legend.position="none",
        legend.background = element_rect(colour = "black"),
        legend.text = element_text(size=8),axis.text.x = element_text(angle = 60, hjust=1),
        legend.title = element_text(size=8),
        axis.title.y = element_text(size=8),
        legend.key.size = unit(.1, "cm"))

ggsave(here("figures","figure_S6_p1_wolegend.png"),width=7,height=2.5)
```


```{r}
# panel 2 with out legend

tmp_analysis_nogroup %>% 
  relabel_program() %>% ggplot() +
  scale_x_continuous(breaks=c(as_date("2016-07-01"),as_date("2017-07-01"),as_date("2018-07-01"),as_date("2019-07-01"),as_date("2020-07-01"),as_date("2021-07-01"),as_date("2022-07-01"),as_date("2023-07-01"),as_date("2024-07-01")),
                     labels = paste("July",2016:2024)) + 

  geom_vline(xintercept=as_date("2020-03-01"),linetype="solid") + 
  geom_vline(xintercept=as_date("2021-07-01"),linetype="dashed") + 
  geom_vline(xintercept=as_date("2021-12-01"),linetype="dotted") + 
  geom_vline(xintercept=as_date("2023-05-01"),linetype="dotdash") + 
  geom_point(aes(x=first_date,y=tele,colour=program)) +
  geom_line(aes(x=first_date,y=tele,colour=program)) + 
  theme_bw() +   
  scale_colour_manual(name="Department",values=c("#FFA47F","#1E99E5")) + 
  theme(legend.position="none",
        legend.background = element_rect(colour = "black"),axis.text.x = element_text(angle = 60, hjust=1),
        legend.text = element_text(size=8),
        legend.title = element_text(size=8),
        axis.title.y = element_text(size=8),
        legend.key.size = unit(.1, "cm")) + 
  ylab("% telehealth visits") + xlab("")
ggsave(here("figures","figure_S6_p2_wolegend.png"),width=7,height=2.5)
```


```{r}
# panel 2 with legend
tmp_analysis_nogroup %>% 
  relabel_program() %>% ggplot() +
  scale_x_continuous(breaks=c(as_date("2016-07-01"),as_date("2017-07-01"),as_date("2018-07-01"),as_date("2019-07-01"),as_date("2020-07-01"),as_date("2021-07-01"),as_date("2022-07-01"),as_date("2023-07-01"),as_date("2024-07-01")),
                     labels = paste("July",2016:2024)) + 

  geom_vline(xintercept=as_date("2020-03-01"),linetype="solid") + 
  geom_vline(xintercept=as_date("2021-07-01"),linetype="dashed") + 
  geom_vline(xintercept=as_date("2021-12-01"),linetype="dotted") + 
  geom_vline(xintercept=as_date("2023-05-01"),linetype="dotdash") + 
  geom_point(aes(x=first_date,y=tele,colour=program)) +
  geom_line(aes(x=first_date,y=tele,colour=program)) + 
  theme_bw() +   
  scale_colour_manual(name="Department",values=c("#FFA47F","#1E99E5")) + 
  theme(legend.position=c(.2,.75),
        legend.background = element_rect(colour = "black"),axis.text.x = element_text(angle = 60, hjust=1),
        legend.text = element_text(size=8),
        legend.title = element_text(size=8),
        axis.title.y = element_text(size=8),
        legend.key.size = unit(.1, "cm")) + 
  ylab("% telehealth visits") + xlab("")
ggsave(here("telehealth","telehealthRR","figures","figure_S6_p2_wlegend.png"),width=7,height=2.5)
```


```{r fig.width=7,fig.height=2.5}
hosp <- readr::read_csv(here("data","externaldata","MDCOVID19_TotalCurrentlyHospitalized_AdultandPediatric_AcuteAndICU_10222024.csv"),show_col_types=FALSE) %>%
  mutate(first_date = lubridate::as_date(ReportDate)) %>% filter(Total!=0)
```


```{r fig.width=7,fig.height=2.5}
# panel 3
hosp %>%
  ggplot() + 
  geom_vline(xintercept=as_date("2020-03-01"),linetype="solid") + 
  geom_vline(xintercept=as_date("2021-07-01"),linetype="dashed") + 
  geom_vline(xintercept=as_date("2021-12-01"),linetype="dotted") + 
  geom_vline(xintercept=as_date("2023-05-01"),linetype="dotdash") + 
  geom_line(aes(x=first_date,y=Total),colour="#810000") + 
  annotate(geom="label",x=as_date("2019-11-01"),y=4500,label="March 2020:\nFirst COVID-19 cases detected in MD,\n State of Emergency in MD is declared",size=2) + 
  annotate(geom="label",x=as_date("2021-5-01"),y=6750,label="July 2021:\n State of Emergency in MD ends,\n Delta variant detected in MD",size=2) + 
  annotate(geom="label",x=as_date("2022-3-01"),y=4500,label="December 2021:\n Omicron variant detected in MD",size=2) +
  
  annotate(geom="label",x=as_date("2023-9-01"),y=2750,label="May 2023:\n National COVID-19 \n Emergency Declaration Ends",size=2) +
  #annotate(geom="label",x=as_date("2020-03-25"),y=2000,label="(Hospitalization data is available from 3/25/20)",size=2) +
 
  theme_bw() +    
  theme(axis.title.y = element_text(size=8),axis.text.x = element_text(angle = 60, hjust=1)) +   scale_y_continuous(limits=c(0,7500))+
  
   scale_x_continuous(breaks=c(as_date("2016-07-01"),as_date("2017-07-01"),as_date("2018-07-01"),as_date("2019-07-01"),as_date("2020-07-01"),as_date("2021-07-01"),as_date("2022-07-01"),as_date("2023-07-01"),as_date("2024-07-01")),
                     labels = paste("July",2016:2024),limits = c(min(tmp_analysis_nogroup$first_date),max(tmp_analysis_nogroup$first_date))) + 

  ylab("Daily COVID-19 Hospital Beds") + xlab("")

ggsave(here("telehealth","telehealthRR","figures","figure_S6_p3.png"),width=7,height=2.5)
```


# Descriptives

## Descriptive tables

### stats on number of visits  per patient 
```{r}
reg_enc_psych %>% count(cohort_id) %>% summarize(median(n),mean(n),min(n),max(n))
reg_enc_primary %>% count(cohort_id) %>% summarize(median(n),mean(n),min(n),max(n))
```

### Sample descriptives stratified by primary versus PSYCH (all time)

```{r descriptives}
var_labels_desc=list(
    "telehealth" = "Telehealth",
    "in_both_cohorts" = "Patients in both primary care and psychiatry cohorts",
    "n_visits" = "Number of visits per patient",
    "any_tele" = "Any Telehealth Appointments",
    "in_pre" = "Patient seen during pre-pandemic",
    "in_post" = "Patient seen during post-pandemic",
    "in_emerging" = "Patient seen during emerging pandemic",
    "in_early" = "Patient seen during early pandemic",
    "in_pre_post" = "Patient seen both pre- and post- pandemic",
    "marital_status"="Marital Status",
    "time_point" = "Time Period",
    "gender"="Gender",
    "race"="Race",
    "ethnicity"="Ethnicity",
    "urban" = "Urbanicity",
    "age_cat"="Age",
    "age_first_visit" = "Age at First Visit",
    "ADI_group" = "Area deprivation index group",
    "adi_nat" = "Area deprivation index National Percentile 2020",
    "employ" = "Employment",
    "mode_payor_cat" = "Insurance",
    "co_anx" = "Comorbid Anxiety",
    "co_sud" = "Comorbid Substance Use Disorder",
    "day" = "Day of Week",
    "daytime_grp" = "Time of Day",
    "county_top" = "County")

var_demo_list=c("in_both_cohorts",
                "ADI_group",
               "gender",
               "age_cat",
               "race",
               "ethnicity",
               "marital_status",
               "employ",
               "mode_payor_cat",
               "urban",
               "co_anx",
               "co_sud",
               "county_top",
               "in_pre",
               "in_emerging",
               "in_early",
               "in_post",
               "in_pre_post",
               "n_visits")

psych_demo <- reg_enc_psych %>% 
  filter(!duplicated(cohort_id)) %>%
  select(all_of(var_demo_list)) %>%
  tbl_summary(label = var_labels_desc,digits = all_categorical() ~ c(0,1))%>%
  modify_header(label ~ "Characteristic",all_stat_cols() ~ "PSYCH all, N={N}")

primary_demo <- reg_enc_primary %>%
  filter(!duplicated(cohort_id)) %>%
  select(all_of(var_demo_list)) %>%
  tbl_summary(label = var_labels_desc,digits = all_categorical() ~ c(0,1)) %>%
  modify_header(label ~ "Characteristic",all_stat_cols() ~ "primary all, N={N}")

psych_pre_demo <- reg_enc_psych %>%
  filter(time_point == "Pre-pandemic") %>%
  filter(!duplicated(cohort_id)) %>%
  select(all_of(var_demo_list)) %>%
  tbl_summary(label = var_labels_desc,digits = all_categorical() ~ c(0,1)) %>%
  modify_header(label ~ "Characteristic",all_stat_cols() ~ "PSYCH pre, N={N}")

psych_post_demo <- reg_enc_psych %>%
  filter(time_point == "Post-pandemic") %>%
  filter(!duplicated(cohort_id)) %>%
  select(all_of(var_demo_list)) %>%
  tbl_summary(label = var_labels_desc,digits = all_categorical() ~ c(0,1)) %>%
  modify_header(label ~ "Characteristic",all_stat_cols() ~ "PSYCH post, N={N}")

primary_pre_demo <- reg_enc_primary %>%
  filter(time_point == "Pre-pandemic") %>%
  filter(!duplicated(cohort_id)) %>%
  select(all_of(var_demo_list)) %>%
  tbl_summary(label = var_labels_desc,digits = all_categorical() ~ c(0,1)) %>%
  modify_header(label ~ "Characteristic",all_stat_cols() ~ "primary pre, N={N}")

primary_post_demo <- reg_enc_primary %>%
  filter(time_point == "Post-pandemic") %>%
  filter(!duplicated(cohort_id)) %>%
  select(all_of(var_demo_list)) %>%
  tbl_summary(label = var_labels_desc,digits = all_categorical() ~ c(0,1)) %>%
  modify_header(label ~ "Characteristic",all_stat_cols() ~ "primary post, N={N}")


psych_emerging_demo <- reg_enc_psych %>%
  filter(time_point == "Emerging-pandemic") %>%
  filter(!duplicated(cohort_id)) %>%
  select(all_of(var_demo_list)) %>%
  tbl_summary(label = var_labels_desc,digits = all_categorical() ~ c(0,1)) %>%
  modify_header(label ~ "Characteristic",all_stat_cols() ~ "Psych emerging, N={N}")


primary_emerging_demo <- reg_enc_primary %>%
  filter(time_point == "Emerging-pandemic") %>%
  filter(!duplicated(cohort_id)) %>%
  select(all_of(var_demo_list)) %>%
  tbl_summary(label = var_labels_desc,digits = all_categorical() ~ c(0,1)) %>%
  modify_header(label ~ "Characteristic",all_stat_cols() ~ "primary emerging, N={N}")


psych_early_demo <- reg_enc_psych %>%
  filter(time_point == "Early-pandemic") %>%
  filter(!duplicated(cohort_id)) %>%
  select(all_of(var_demo_list)) %>%
  tbl_summary(label = var_labels_desc,digits = all_categorical() ~ c(0,1))  %>%
  modify_header(label ~ "Characteristic",all_stat_cols() ~ "PSYCH early, N={N}")


primary_early_demo <- reg_enc_primary %>%
  filter(time_point == "Early-pandemic") %>%
  filter(!duplicated(cohort_id)) %>%
  select(all_of(var_demo_list)) %>%
  tbl_summary(label = var_labels_desc,digits = all_categorical() ~ c(0,1)) %>%
  modify_header(label ~ "Characteristic",all_stat_cols() ~ "primary early, N={N}")


t1 <- tbl_merge(list(primary_demo,primary_pre_demo, primary_emerging_demo,primary_early_demo,  primary_post_demo,psych_demo,psych_pre_demo,psych_emerging_demo,psych_early_demo,  psych_post_demo))
t1 

```

### median number of visits per patient for each time period (include in Table 1)

```{r descriptives_stt} 

reg_enc_combo %>% group_by(program,cohort_id) %>% summarize(n_visits=n()) %>% group_by(program) %>% summarize(n_pats = n(),median_visits = median(n_visits),q1=quantile(n_visits,prob=.25),q3=quantile(n_visits,prob=.75))


reg_enc_combo %>% group_by(program,cohort_id,time_point) %>% summarize(n_visits=n()) %>% group_by(program,time_point) %>% summarize(n_pats = n(),median_visits = median(n_visits),q1=quantile(n_visits,prob=.25),q3=quantile(n_visits,prob=.75))
```


### telehealth numbers visits (also in Table 1)

```{r}

var_enc_list=c("telehealth")

psych_enc <- reg_enc_psych %>%
  select(all_of(var_enc_list)) %>%
  tbl_summary(label = var_labels_desc,digits = all_categorical() ~ c(0,1)) %>%
  modify_header(label ~ "Characteristic",all_stat_cols() ~ "PSYCH all, N={N}")

primary_enc <- reg_enc_primary %>%
  select(all_of(var_enc_list)) %>%
  tbl_summary(label = var_labels_desc,digits = all_categorical() ~ c(0,1)) %>%
  modify_header(label ~ "Characteristic",all_stat_cols() ~ "primary all, N={N}")


psych_pre_enc <- reg_enc_psych %>%
  filter(time_point == "Pre-pandemic") %>%
  select(all_of(var_enc_list)) %>%
  tbl_summary(label = var_labels_desc,digits = all_categorical() ~ c(0,1)) %>%
  modify_header(label ~ "Characteristic",all_stat_cols() ~ "PSYCH pre, N={N}")

psych_early_enc <- reg_enc_psych %>%
  filter(time_point == "Early-pandemic") %>%
  select(all_of(var_enc_list)) %>%
  tbl_summary(label = var_labels_desc,digits = all_categorical() ~ c(0,1)) %>%
  modify_header(label ~ "Characteristic",all_stat_cols() ~ "PSYCH early, N={N}")


psych_post_enc <- reg_enc_psych %>%
  filter(time_point == "Post-pandemic") %>%
  select(all_of(var_enc_list)) %>%
  tbl_summary(label = var_labels_desc,digits = all_categorical() ~ c(0,1)) %>%
  modify_header(label ~ "Characteristic",all_stat_cols() ~ "PSYCH post, N={N}")

psych_emerging_enc <- reg_enc_psych %>%
  filter(time_point == "Emerging-pandemic") %>%
  select(all_of(var_enc_list)) %>%
  tbl_summary(label = var_labels_desc,digits = all_categorical() ~ c(0,1)) %>%  modify_header(label ~ "Characteristic",all_stat_cols() ~ "PSYCH emerging, N={N}")

primary_pre_enc <- reg_enc_primary %>%
  filter(time_point == "Pre-pandemic") %>%
  select(all_of(var_enc_list)) %>%
  tbl_summary(label = var_labels_desc,digits = all_categorical() ~ c(0,1))%>%
  modify_header(label ~ "Characteristic",all_stat_cols() ~ "primary pre, N={N}")

primary_early_enc <- reg_enc_primary %>%
  filter(time_point == "Early-pandemic") %>%
  select(all_of(var_enc_list)) %>%
  tbl_summary(label = var_labels_desc,digits = all_categorical() ~ c(0,1))%>%
  modify_header(label ~ "Characteristic",all_stat_cols() ~ "primary early, N={N}")

primary_post_enc <- reg_enc_primary %>%
  filter(time_point == "Post-pandemic") %>%
  select(all_of(var_enc_list)) %>%
  tbl_summary(label = var_labels_desc,digits = all_categorical() ~ c(0,1)) %>%
  modify_header(label ~ "Characteristic",all_stat_cols() ~ "primary post, N={N}")

primary_emerging_enc <- reg_enc_primary %>%
  filter(time_point == "Emerging-pandemic") %>%
  select(all_of(var_enc_list)) %>%
  tbl_summary(label = var_labels_desc,digits = all_categorical() ~ c(0,1)) %>%
  modify_header(label ~ "Characteristic",all_stat_cols() ~ "primary emerging, N={N}")

tenc <- tbl_merge(list(primary_enc,primary_pre_enc, primary_emerging_enc,primary_early_enc, primary_post_enc,psych_enc,psych_pre_enc,psych_emerging_enc, psych_early_enc, psych_post_enc))
tenc
```

## Table S1

```{r}
psych_ADI <- reg_enc_psych %>% 
  filter(!duplicated(cohort_id)) %>%
  select(all_of(var_demo_list)) %>%
  tbl_summary(label = var_labels_desc,digits = all_categorical() ~ c(0,1),by="ADI_group") 

primary_ADI <- reg_enc_primary %>%
  filter(!duplicated(cohort_id)) %>%
  select(all_of(var_demo_list)) %>%
  tbl_summary(label = var_labels_desc,digits = all_categorical() ~ c(0,1),by="ADI_group") 

final_ADI_tbl <- tbl_merge(list(primary_ADI,psych_ADI), tab_spanner = c("**Mental health care in primary care**", "**Mental health care in psychiatry**"))

final_ADI_tbl

```

# RQ-1 Analysis
```{r}
tmp_analysis <- rbind(reg_enc_primary,reg_enc_psych) %>% 
  group_by(contact_date,program,month,year,ADI_group,time_point, t_weeks,day) %>% 
  filter(!weekend) %>%
  summarize(n =n()) %>% 
  ungroup() %>% filter(time_point !="Emerging-pandemic",time_point !="Early-pandemic") %>%
  mutate(t_weeks = t_weeks/52)
```

```{r}
tmp_analysis %>% group_by(time_point) %>% summarize(min(t_weeks),max(t_weeks))
```


```{r}
tmp_primary <- filter(tmp_analysis,program=="primary")%>% droplevels()

tmp_psych <- filter(tmp_analysis,program=="PSYCH") %>% droplevels()

tmp_psych$month <- relevel(factor(tmp_psych$month,ordered=FALSE), ref = "Jul")

tmp_primary$month <- relevel(factor(tmp_primary$month,ordered=FALSE), ref = "Jul")


formula <- as.formula("n ~ time_point*ADI_group*I(t_weeks-5) + factor(day,ordered=FALSE)+ factor(month,ordered=FALSE)")
 
m0_psych_nb<- MASS::glm.nb(formula,tmp_psych)
m0_primary_nb<- MASS::glm.nb(formula,tmp_primary)

m0_primary_nb_table <- tbl_regression(m0_primary_nb,exponentiate=TRUE,pvalue_fun = label_style_pvalue(digits = 2))
m0_psych_nb_table <- tbl_regression(m0_psych_nb,exponentiate=TRUE,pvalue_fun = label_style_pvalue(digits = 2))

order <- c("time_point", "I(t_weeks - 5)", "ADI_group","time_point:I(t_weeks - 5)", "time_point:ADI_group",  "ADI_group:I(t_weeks - 5)", "time_point:ADI_group:I(t_weeks - 5)","factor(day, ordered = FALSE)", "factor(month, ordered = FALSE)") 

m0_primary_nb_table_out <- m0_primary_nb_table$table_body %>%
  select(variable,label,estimate,conf.low,conf.high,p.value) %>%
  mutate(variable=factor(variable,levels=order) ) %>%
  arrange(variable) %>% mutate(estimate = round(estimate,2),
                             conf.low = round(conf.low,2),
                             conf.high = round(conf.high,2),
                            pround2 = round(p.value,2))

m0_psych_nb_table_out <- m0_psych_nb_table$table_body %>%
  select(variable,label,estimate,conf.low,conf.high,p.value) %>%
  mutate(variable=factor(variable,levels=order) ) %>% 
  arrange(variable) %>% mutate(estimate = round(estimate,2),
                             conf.low = round(conf.low,2),
                             conf.high = round(conf.high,2),
                            pround2 = round(p.value,2))
```

### two parts of table S2
```{r}
m0_primary_nb_table_out
m0_psych_nb_table_out

```

### linear combinations (table S3)
```{r}
marginal_means_primary_nb <-avg_predictions(m0_primary_nb, by=c("ADI_group","time_point")) %>%
  mutate(program="Mental health in primary care",
         estimating = "mean (average number of visits-per-day during time period)",
         estimate_pretty = style_ratio(estimate,digits=2),
         conf_pretty = paste0(style_ratio(conf.low,digits=2),", ",style_ratio(conf.high,digits=2)),
        p.value_rounded = style_pvalue(p.value,digits = 2)) %>% 
  relocate(program,estimating,time_point,ADI_group,estimate_pretty,conf_pretty,p.value_rounded,estimate,conf.low,conf.high,p.value) %>% as_tibble()

marginal_means_psych_nb <-avg_predictions(m0_psych_nb, by=c("ADI_group","time_point")) %>%
  mutate(program="Mental health in psychiatry",
  estimating = "mean (average number of visits-per-day during time period)",
        estimate_pretty = style_ratio(estimate,digits=2),
         conf_pretty = paste0(style_ratio(conf.low,digits=2),", ",style_ratio(conf.high,digits=2)),
        p.value_rounded = style_pvalue(p.value,digits = 2)) %>% 
  relocate(program,estimating,time_point,ADI_group,estimate_pretty,conf_pretty,p.value_rounded,estimate,conf.low,conf.high,p.value) %>% as_tibble()

marginal_slopes_primary_nb <-avg_comparisons(m0_primary_nb, variables = "t_weeks",by=c("time_point","ADI_group"),comparison="lnratioavg",transform="exp") %>%
  mutate(program="Mental health in primary care",
         estimating = "slope (change in visits-per-day over 1 year in time period)",
         perc_change = round((estimate - 1)*100,1),
         perc_change_conf.low = round((conf.low - 1)*100,1),
         perc_change_conf.high = round((conf.high -1)*100,1),p.value_rounded =style_pvalue(p.value,digits = 2),
          conf_pretty = paste0(perc_change_conf.low,", ",perc_change_conf.high)) %>% 
  relocate(program,estimating,time_point,ADI_group,perc_change,conf_pretty,perc_change_conf.low,perc_change_conf.high,p.value_rounded,estimate,conf.low,conf.high,p.value) %>% as_tibble()

marginal_slopes_psych_nb <-avg_comparisons(m0_psych_nb,  variables = "t_weeks",by=c("time_point","ADI_group"),comparison="lnratioavg",transform="exp")%>%
  mutate(program="Mental health in psychiatry",
         estimating = "slope (change in visits-per-day over 1 year in time period)",
         perc_change = round((estimate - 1)*100,1),
         perc_change_conf.low = round((conf.low - 1)*100,1),
         perc_change_conf.high = round((conf.high -1)*100,1),p.value_rounded = style_pvalue(p.value,digits = 2),conf_pretty = paste0(perc_change_conf.low,", ",perc_change_conf.high)) %>% 
  relocate(program,estimating,time_point,ADI_group,perc_change,conf_pretty,perc_change_conf.low,perc_change_conf.high,p.value_rounded,estimate,conf.low,conf.high,p.value)  %>% as_tibble()

marginal_means_primary_nb
marginal_means_psych_nb
marginal_slopes_primary_nb
marginal_slopes_psych_nb
```

### supplemental analysis (no linear trends, just pre-post by group comparison)
```{r}
formula_base <- as.formula("n ~ time_point*ADI_group")
 
m0_psych_nb_base<- MASS::glm.nb(formula_base,tmp_psych)
m0_primary_nb_base<- MASS::glm.nb(formula_base,tmp_primary)

m0_primary_nb_table_base <- tbl_regression(m0_primary_nb_base,exponentiate=TRUE,pvalue_fun = label_style_pvalue(digits = 2))
m0_psych_nb_table_base <- tbl_regression(m0_psych_nb_base,exponentiate=TRUE,pvalue_fun = label_style_pvalue(digits = 2))
m0_primary_nb_table_base
m0_psych_nb_table_base

pre_post_compare_primary_base <- avg_comparisons(m0_primary_nb_base, variables = "time_point",by=c("ADI_group"),comparison="lnratioavg",transform="exp")

pre_post_compare_psych_base <- avg_comparisons(m0_psych_nb_base, variables = "time_point",by=c("ADI_group"),comparison="lnratioavg",transform="exp")

pre_post_compare_primary_base
pre_post_compare_psych_base
```


```{r}
writexl::write_xlsx(list(table_1_p1_visit_level = gtsummary::as_tibble(tenc),
                         table_1_p2_pat_level= gtsummary::as_tibble(t1),
                         table_S1_by_ADI=gtsummary::as_tibble(final_ADI_tbl),
                         
                         table_S2_primarycare = gtsummary::as_tibble(m0_primary_nb_table_out),
                         table_S2_psychiatry = gtsummary::as_tibble(m0_psych_nb_table_out),
                                             
                         base_primarycare = gtsummary::as_tibble(m0_primary_nb_table_base),
                         base_psychiatry = gtsummary::as_tibble(m0_psych_nb_table_base),
                         
                         tableS3_part1_means_primarycare = marginal_means_primary_nb,
                         tableS3_part2_slopes_primarycare = marginal_slopes_primary_nb,
                         tableS3_p3_oneyear_primarycare = one_year_compare_primary, 
                         tableS3_p4_pre_post_primarycare = pre_post_compare_primary,
                        
                         
                         tableS3_part1_means_psychiatry =marginal_means_psych_nb,
                         tableS3_part2_slopes_psychiatry =marginal_slopes_psych_nb,
                         tableS3_p3_oneyear_psych = one_year_compare_psych, 
                         tableS3_p4_pre_post_psych = pre_post_compare_psych),
                    path= here("RQ1-tables.xlsx"))
```



